### 2021-02-26 yfcc100m/paper/vldb2021/design.tex
Minor fixes

VDMS implements a the typical client-server architecture that handles
client queries transactionally and concurrently, as most of the most common
relational and non-relational database systems
\cite{mysql, postgresql, chang2008bigtable}.
The main difference between other data management systems and VDMS is that
it goes beyond the typical supported data types (string, integers, floats,
blobs, JSON-documents, etc), recognizing visual entities (image, videos,
feature vectors, etc) as first class citizens.
VDMS API enable users to insert, index, process, and query visual data,
as well as provides full support for inserting, indexing and querying
user-defined metadata.
Users interact with both metadata and visual data using a unified API,
in a transactional manner.

VDMS provides a property \textit{graph model} abstraction with the traditional
atomicity, consistency, isolation, and durability
(ACID) properties expected from databases.
This is, users interact with their objects (metadata, images, videos, etc),
as if these objects were in a connected graph.
Graph represents an easier abstraction to model complex problems,
making it very suitable for the data and access patterns shown by visual metadata,
which can be easily mapped into application-level abstractions by
developers~\cite{tao}.
For instance, abstractions like \textit{BoundingBoxes} associated to
images or videos can be easily represented using nodes and edges in a graph.
These are the reasons the team chose a graph database over a
relational database as the metadata management for the implementation of VDMS.

VDMS API provides a mechanism to insert and connect Images, Videos, BoundingBoxes,
Frames, and Descriptors (feature vectors), together with any metadata
associated with the objects.
Each object is a node in the graph.
The metadata associated to each object is modeled as
"properties" of the node in the graph.
Users can query, filter, and retrieve this objects based
on its metadata (the "properties" associated with them).
VDMS does not simply treat these objects as binary blobs of data, but rather
understands them and the type of processing that is common for them,
providing the ability to run processing on-the-fly, both at insertion and
retrieval.
Users can, for instance, retrieve a resized version of an image,
or retrieve only a clip from a large video.
This is one of the main differentiation aspect when compared to other
database systems, including relational databases.
In a relational database, for instance, one can query \textit{and compute}
over values in a column only for basic data types (strings, integers, floats),
and over the abstractions the relational model supports (tables, columns).
A SQL query can then retrieve the \textit{computed average} "salary" of all
the employees in a company (stored using the table/columns abstraction),
but cannot perform any computation over data stored as a blob or binary object.
VDMS, because it recognized the nature of visual objects by design,
provides the ability to \textit{compute} on these visual objects
(image, videos, etc).

@@

By providing the ability to store visual data objects together with
application-defined entities and its properties, VDMS can manage
all the data the application need behind a single, unified API.
This is in contrast with current applications that rely on a combination
of multiple systems and APIs to access different portions of the data they need
\cite{tao, sculley2015hidden, mayer2020scalable, sculley2015hidden}.

@@

One of the most important differentiating aspects of VDMS is its API.
VDMS is unique in recognizing visual entities (i.e., images, videos, etc)
as first class citizens.
Thus, VDMS' API revolves around visual data operations and retrieval.
VDMS API is easy to use and explicitly pre-defines certain
primitives associated with metadata, images, videos, and feature vectors.
Authors have paid particular attention to hide the complexities of our internal
implementation and up-level the API to a JSON-based API,
which is very popular across various application domains.
By defining a new JSON-based API, there is a trade-off between expressiveness
(compared to well-established query languages like SPARQL, Gremlim, or even SQL)
and the ability to natively support visual data operations.
However, we believe it is possible for our API to achieve similar levels of
expressiveness compared to more mature query languages over time.

Listing~\ref{addimageandautotag} shows a sample query for inserting an image
and its associated metadata to VDMS.
In this example, the transaction inserts an application-defined "Entity" of
the class "autotag", with the "name" property being \textit{alligator},
inserts an Image with its "latitude" and "longitude",
stores the image as a JPG (VDMS will transcode if needed),
and creates a connection between the Image and the Entity, with a "prob"
property (which indicates the probability of that image
containing an object of type  \textit{alligator}) with a specific value.
Note that no schema needs to be defined in advance.
The Entity of class "autotag", with its properties, are declared and added
at insertion time, without any need to define a schema of objects
(and its properties) before hand.

@@

Listing~\ref{findimagegeo} shows another sample query, in this case, for retrieval.
In this particular example, the transaction retrieves all the images
of \textit{alligators} with probability higher than 0.66,
filter by latitude and longitude within 1 degree,
apply a resize operation to make the images 224x224,
rotate the images 45.34 degrees, and return the images as "png" files.
It is important to note how the API natively supports basic building
blocks of visual data processing, like resize, rotation, or transcoding
(changing output formats and encodings).
The API allows interaction with metadata, images, videos,
and feature vectors in a similar fashion, and it is fully
documented on the project's Github
wiki~\footnote{https://github.com/IntelLabs/vdms/wiki/API-Description}.

\begin{listing}[ht!]
\begin{minted}[frame=single,
              framesep=3mm,
              linenos=true,
              xleftmargin=21pt,
              tabsize=4]{js}
"FindEntity"{
    "class": "autotag",
    "constraints": {
        "name": ["==", "alligator"]
    }
    "_ref" : 1
},
"FindImage":{
    "format": "png",
    "link": {
        "ref":1,
        "constraints": {
            "prob": [">=", 0.66]
        }
    },
    "constraints": {
        "latitude": [">=", 36.23433,
                     "<=", 38.23433]
        "longitude":[">=", -114.80666,
                     "<=", -116.80666]
    },
    "operations": [{
            "type": "resize",
            "height": 224,
            "width":  224,
        }, {
            "type": "rotate",
            "angle": 45.34
    }]
}

@@

The Graph Engine used in VDMS is the
\textit{Persistent Memory Graph Database} (PMGD).
PMGD was developed at Intel Labs, and was designed and optimized for
persistent memory technologies like Intel Optane~\cite{IntelXPoint15}, which
promise storage providing nearly the speed of DRAM and the
durability of block-oriented storage.
PMGD is also fully open-sourced \footnote{https://github.com/IntelLabs/pmgd}.
PMGD implements an in-persistent-memory graph database, PMGD, optimized
to run on a platform equipped with persistent memory, but
also provides the option to run on SSD and ext4 filesystem
using OS support to provide transactional guarantees (through \textit{msync}).
PMGD provides a property graph model of data storage with the traditional
atomicity, consistency, isolation, and durability
(ACID) properties expected from databases.
PMGD comes as a C++ library that VDMS uses for its internal query engine.
Because PMGD already supports most of the graph abstractions we wanted
to see on VDMS API, it was our preferred option.
Also, our internal evaluation shows that PMGD is faster
than other graph databases, including Neo4j\cite{miller2013graph}.
PMGD performance evaluation will be published separately.

@@

The Visual Compute Module was designed and implemented to provide
an internal abstraction layer for interacting with visual data.
It enables simple visual data handling and processing
(i.e., basic building block operations like crop, resize, etc).
For traditional formats (jpg, png, tiff, mp4, etc.),
the interface is an abstraction layer over OpenCV.
However, it also provides a way to use novel formats that are
better suited for visual analytics: a novel, array-based lossless image format.
This format is built on the array data manager TileDB~\cite{TileDB} and
is well suited for images that are used in visual analytics.
Note that, even if VDMS currently provides support for array-based lossless
image format, we do not use it as part of this evaluation.
This work focuses on a more direct comparison with a combination of alternatives,
and thus we use the traditional format (jpg and png).
The performance comparison between the array-based lossless image format
we develop and other similar formats (like PNG or TIFF) are outside of the
scope of this evaluation.

VDMS also provides full support for video storage and operations,
in a similar way it does for images.
This includes support for encoding, decoding, and transcoding of
\textit{mp4}, \textit{avi}, and \textit{mov} containers,
as well as support for \textit{xvid}, \textit{H.263} and \textit{H.264} encoders.
This is supported through the use of either OpenCV~\cite{opencv}
or \textit{libffmpeg}\cite{ffmpeg}, or both.
All operations supported for images in VDMS are also supported at the
video and frame level of the API.
On top of that, there are a number of video-specific operations that
are supported, such as the interval operations,
enabling users to retrieve clips at different
frames-per-second (FPS) versions of the video.

Another key differentiating factor of VDMS is that it allows the creation of
indexes for high-dimensional feature vectors and the insertion of
these feature vectors associated with entities, images, and/or videos.
Feature vectors are intermediate results of various machine
learning or computer vision algorithms when run on visual data.
Feature vectors are also known as \textit{descriptors}
or \textit{visual descriptors}.
We use these terms interchangeably.
These descriptors can be classified, labeled, and used to build search indexes.
Feature Vectors support is provided through our implementation based
on high-dimensional sparse arrays, also using array-based approaches.
In addition, the Visual Compute Library provides a wrapper
for another high-dimensional index implementation,
Facebook's Faiss~\cite{faiss}.

@@

The client library implements TCP/IP based connectors to the VDMS Server,
similar to most databases\cite{memsql, mysql, postgresql}.
Users can connect to VDMS and implement queries using VDMS' API
by defining JSON commands conforming to the query protocol we have defined.
The client library provides a simple method that
accepts a JSON string and an array or vector of blobs.
Internally, the library wraps the query string and blobs using
Google Protobufs \cite{protobufs} and sends it to the VDMS server.
It also receives a similarly formed response from VDMS
and returns it to the client.
The responses require JSON parsing on the client
side for the metadata string that indicates how to interpret the blobs field.
Currently, client libraries are implemented for Python and C++ client.
The client libraries are lightweight, as they simply implement the communication
protocol between the client and the server.
This makes it easier for developers to implement similar client libraries using
any other programming language of their choice.

>>>>

VDMS implements the typical client-server architecture that handles
client queries transactionally and concurrently, similar to what most common
relational and non-relational database systems
\cite{mysql, postgresql, chang2008bigtable} use.
The main difference between other data management systems and VDMS is that
it goes beyond the typical supported data types (string, integers, floats,
blobs, JSON-documents, etc), recognizing visual entities (image, videos,
feature vectors, etc) as first class citizens.
VDMS API enable users to insert, index, process, and query visual data,
as well as provides full support for inserting, indexing and querying
user-defined metadata.
Users interact with both metadata and visual data using a unified API,
in a transactional manner.

VDMS provides a \textit{graph model} abstraction with the traditional
atomicity, consistency, isolation, and durability
(ACID) properties expected from databases.
This is, users interact with their objects (metadata, images, videos, etc),
as if these objects were in a connected graph.
Graph represents an easier abstraction to model complex problems,
making it very suitable for the data and access patterns shown by visual metadata,
which can be easily mapped into application-level abstractions by
developers~\cite{tao}.
For instance, abstractions like \textit{BoundingBoxes} associated to
images or videos can be easily represented using nodes and edges in a graph.
This is the main reason why the team chose a graph model over a
relational one for the implementation of VDMS API.

VDMS API provides a mechanism to insert and connect Images, Videos, BoundingBoxes,
Frames, and Descriptors (feature vectors), together with any metadata
associated with the objects.
Each object is a node in the graph.
The information associated to each visual object (image, video, etc) is modeled as
"properties" of the node in the graph.
Users can query, filter, and retrieve this objects based on its properties.
VDMS does not simply treat these objects as binary blobs of data, but rather
understands them and the type of processing that is common for them,
providing the ability to run processing on-the-fly, both at insertion and
retrieval time.
Users can, for instance, retrieve a resized version of an image,
or retrieve only a clip from a large video.
This is one of the main differentiation aspect when compared to other
database systems, including relational databases.
In a relational database, for instance, one can query \textit{and compute}
over values in a column only for basic data types (strings, integers, floats),
and over the abstractions the relational model supports (tables, columns).
This is, a SQL query can retrieve the \textit{computed average} "salary" of all
the employees in a company (stored using the table/columns abstraction),
but cannot perform any computation over data stored as a blob or binary object.
VDMS, because it recognized the nature of visual objects by design,
provides the ability to \textit{compute} on these visual objects
(image, videos, etc).

@@

By providing the ability to store visual data objects together with
application-defined entities and its properties, VDMS can manage
all the data the application need behind a single, unified API.
This is in contrast with current applications that rely on a combination
of multiple data management systems and APIs to access different
portions of the data they need
\cite{tao, sculley2015hidden, mayer2020scalable, sculley2015hidden}.

@@

One of the most important differentiating aspects of VDMS is its API.
VDMS is unique in recognizing visual entities (i.e., images, videos, etc)
as first class citizens.
Thus, VDMS' API revolves around visual data operations and retrieval,
but at the same time, enables applications to store any other
application-specific metadata.
VDMS API is easy to use and explicitly pre-defines certain
primitives associated with metadata, images, videos, and feature vectors.
Authors have paid particular attention to hide the complexities of our internal
implementation and up-level the API to a JSON-based API,
which is very popular across various application domains.
By defining a new JSON-based API, there is a trade-off between expressiveness
(compared to well-established query languages like SPARQL, Gremlim, or even SQL)
and the ability to natively support visual data operations.
However, we believe it is possible for our API to achieve similar levels of
expressiveness compared to more mature query languages over time.

Listing~\ref{addimageandautotag} shows a sample query for inserting
an image, properties associated with the image, and metadata to VDMS.
The metadata, in this case, is the information about the "autotag",
which is provided by the dataset in this specific
\textit{image-search} application.
In this example, the transaction inserts an application-defined "Entity" of
the class "autotag", with the "name" property being \textit{alligator},
inserts an Image with its "latitude" and "longitude",
stores the image as a JPG (VDMS will transcode if needed),
and creates a connection between the Image and the Entity, with a "prob"
property (which indicates the probability of that image
containing an object of type  \textit{alligator}) with a specific value.
This query is performed \textit{transactionally}: either all the image, metadata
and connection are inserted, or none of them are and an error is returned.
Note that no schema needs to be defined in advance.
The Entity of class "autotag", with its properties, are declared and added
at insertion time, without any need to define a schema of objects
(and its properties) before hand.

@@

Listing~\ref{findimagegeo} shows another sample query, in this case, for retrieval.
In this particular example, the transaction retrieves all the images
of \textit{alligators} with probability higher than 0.66,
filter by latitude and longitude within 1 degree,
apply a resize operation to make the images 224x224,
rotate the images 45.34 degrees, and return the images as "png" files.
It is important to note how the API natively supports basic building
blocks of visual data processing, like resize, rotation, or transcoding
(changing output formats and encodings).
The API allows interaction with metadata, images, videos, bounding boxes,
frames, feature vectors, and more in a similar fashion,
and it is fully documented on the project's Github
wiki~\footnote{https://github.com/IntelLabs/vdms/wiki/API-Description}.

\begin{listing}[ht!]
\begin{minted}[frame=single,
              framesep=3mm,
              linenos=true,
              xleftmargin=21pt,
              tabsize=4]{js}
"FindEntity"{
    "class": "autotag",
    "constraints": {
        "name": ["==", "alligator"]
    }
    "_ref" : 1
},
"FindImage":{
    "format": "png",
    "constraints": {
        "latitude": [">=", 36.23433,
                     "<=", 38.23433]
        "longitude":[">=", -114.80666,
                     "<=", -116.80666]
    },
    "operations": [{
            "type": "resize",
            "height": 224,
            "width":  224,
        }, {
            "type": "rotate",
            "angle": 45.34
    }],
    "link": {
        "ref":1,
        "constraints": {
            "prob": [">=", 0.66]
        }
    }
}

@@

The Graph Engine used in VDMS is the
\textit{Persistent Memory Graph Database} (PMGD).
PMGD was developed at Intel Labs, and was designed and optimized for
persistent memory technologies like Intel Optane~\cite{IntelXPoint15}, which
promise storage providing nearly the speed of DRAM and the
durability of block-oriented storage.
PMGD is also fully open-sourced \footnote{https://github.com/IntelLabs/pmgd}.
PMGD implements an in-persistent-memory graph database optimized
to run on a platform equipped with persistent memory, but
also provides the option to run on SSD and ext4 filesystem
using OS support to provide transactional guarantees (through \textit{msync}).
PMGD provides a property graph model of data storage with the traditional
atomicity, consistency, isolation, and durability
(ACID) properties expected from databases.
PMGD comes as a C++ library that VDMS uses for implementing its query engine.
Because PMGD already supports most of the graph abstractions we wanted
to see on VDMS API, it was our preferred option.
Also, our internal evaluation shows that PMGD is faster
than other graph databases, including Neo4j\cite{miller2013graph}.
PMGD performance evaluation will be published separately.

@@

The Visual Compute Module was designed and implemented to provide
an internal abstraction layer for interacting with visual data.
It enables the query engines to coordinate and perform
visual data handling and processing
(i.e., basic building block operations like crop, resize, etc for images/videos
and k-nearest neighbor search for feature vectors),
shown in Figure \ref{fig:arch}.
For traditional formats (jpg, png, tiff, mp4, etc.),
the interface is an abstraction layer over OpenCV.
However, it also provides a way to use novel formats that are
better suited for visual analytics: a novel, array-based lossless image format.
This format is built on the array data manager TileDB~\cite{TileDB} and
is well suited for images that are used in visual analytics.
Note that, even if VDMS currently provides support for array-based lossless
image format, we do not use it as part of this evaluation.
This work focuses on a more direct comparison with a combination of alternatives,
and thus we use the traditional format (jpg and png).
The performance comparison between the array-based lossless image format
we develop and other similar formats (like png or tiff) are outside of the
scope of this evaluation.

VDMS also provides full support for video storage and operations,
in a similar way it does for images.
This includes support for encoding, decoding, and transcoding of
\textit{mp4}, \textit{avi}, and \textit{mov} containers,
as well as support for \textit{xvid}, \textit{H.263} and \textit{H.264} encoders.
This is supported through the use of either OpenCV~\cite{opencv}
or \textit{libffmpeg}\cite{ffmpeg}, or both,
plus additional implementation to support fast random access to video frames.
All operations supported for images in VDMS are also supported at the
video and frame level of the API.
On top of that, there are a number of video-specific operations that
are supported, such as the interval operations,
enabling users to retrieve clips at different
frames-per-second (fps) versions of the video.

Another key differentiating factor of VDMS is that it allows the creation of
indexes for high-dimensional feature vectors and the insertion of
these feature vectors associated with entities, images, and/or videos.
Feature vectors are intermediate results of various machine
learning or computer vision algorithms when run on visual data.
Feature vectors are also known as \textit{descriptors}
or \textit{visual descriptors}.
We use these terms interchangeably.
These descriptors can be classified, labeled, and used to build search indexes.
Feature Vectors support is provided through our implementation based
on high-dimensional sparse arrays, also using array-based approaches.
In addition, the Visual Compute Library provides a wrapper
for another high-dimensional index implementation,
Facebook's Faiss~\cite{faiss}.
Users can, through our API, use different indexing techniques
for feature vectors, depending on their application's need.

@@

The client library implements TCP/IP based connectors to the VDMS Server,
similar to most databases\cite{memsql, mysql, postgresql}.
Users can connect to VDMS and run queries using VDMS' API
by defining a transaction using JSON objects.
The client library provides a simple method that
accepts a JSON string and an array or vector of blobs.
Internally, the library wraps the query string and blobs using
Google Protobufs \cite{protobufs} and sends it to the VDMS server.
It also receives a similarly formed response from VDMS
and returns it to the client.
Currently, client libraries are implemented for Python and C++ client.
The client libraries are lightweight, as they simply implement the communication
protocol between the client and the server.
This makes it easier for developers to implement similar client libraries using
any other programming language of their choice.

### 2021-02-26 yfcc100m/paper/vldb2021/eval.tex
Minor fixes

We have used the YFCC100M dataset to evaluate different aspects of our system.
We use the images in the dataset and its associated metadata to implement
an image-search engine based on properties associated with those images.
This is a very common use-case we have encountered when building
applications such as smart-retail, sports applications, and video summarization.
For these type of applications, the starting point is usually
a large set of data that must be curated
before proceeding with the data processing (such as neural network training).
In order to evaluate the different aspects of the performance on the
image search, we have built different baselines following the methodology
used in the industry, and following what we have done in the past
in order to solve the data search problem,
which we describe in Section \ref{images}.
% We have also included performance evaluation on the Video and Feature Vectors
% functionality in the appendix for further reference, together with a
% description of the key aspects of those functionalities.

@@

The Yahoo! Flickr Creative Commons 100m (YFCC100M) dataset is a large
collection of 100 million public Flickr media objects created to provide free,
shareable multimedia data for research.
This dataset contains approximately 99.2 million images and 0.8 million videos
with metadata characterized by 25 fields such as the unique identifier, userid,
date the media was taken/uploaded, location in longitude/latitude coordinates,
device type the media was captured, URL to download the media object,
and the Creative Commons license type information.
The YFCC100M dataset also contains \textit{autotags}
provided as a set of comma-separated concepts such as people, scenery, objects,
and animals from 1,570 trained machine learning classifiers ~\cite{Thomee_2016}.
Together with each \textit{autotags}, there is a probability associated with
each tag to indicate certainty of the classification.
This is, an image can have the \textit{autotags} "people", "person", "party",
"outdoor", and each \textit{autotag} assigned will be accompanied by a
probability of that \textit{autotags} being present in that image/frame.
We have also used feature vectors generated for every image and first frame
of every video \cite{features} to implement a similarity search.
Given that there is no standard benchmark oriented towards visual data queries,
we have built a series of queries to filter this dataset that is modeled after
our internal use cases for many of the mentioned applications we have worked
with.

@@

Given that there are no other open-source systems that provide similar
functionality and interfaces as VDMS (i.e., transactionally dealing with images
and metadata behind a single interface), we have implemented two equivalent
visual data management system as baselines, comprised of a combination of
widely available, off-the-shelf components.
The first baseline uses MySQL Server 5.7 (for storing metadata),
Apache Web Server 2.4.18 (as interface for image access), and
OpenCV 3.3 (to provide pre-processing operations on images).
The other baseline system replaces the MySQL Server 5.7 with a
most advanced open-source relational database, PostgreSQL 9.5 \cite{postgresql}.
We decided to use relational databases in the baseline systems instead of
non-relational databases because of their maturity, and also because it is
the most common tools used for storing and querying metadata
~\cite{Jatana2012, li_2019}:

\begin{itemize}
    \item Relational databases support atomicity, consistency, isolation,
    and durability (ACID) while non-relational may compromise some ACID properties.
    \item The YFCC100m data is clearly structured.
    \item We need to efficiently collate and return metadata records.
\end{itemize}

The baseline implementations only partially replicate the functionalities
that VDMS offers when it comes to image and metadata handling, and it was
built for the purpose of an ad-hoc image search implementation.
The baseline implementations are based upon internal tools used for ML-based
pipelines for media, the common practice in the industry \cite{haystack, tao}.
We have implemented a set of client-side applications that take care
of retrieving the components from the different systems, and applies
pre-processing operations when needed.

For all our experiments, we use two servers with Ubuntu 20.04,
one hosting a VDMS server and another hosting the baseline implementations.
Both servers have a dual-socket Intel\textsuperscript{\textregistered}
Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
each CPU with 28 physical cores with hyper-threading enabled,
for a total of 112 logical cores per server.
The server hosting MySQL and PostgreSQL has 256GB of DDR4 DRAM,
while the server hosting VDMS has 194GB of DDR4 DRAM.
We decided to run the VDMS server in the machine with less DRAM to make
sure MySQL and PostgreSQL had no disadvantage. Previous evaluation
indicated smaller footprint in the case of VDMS when
compared to similar baselines based on MySQL.
To build both MySQL and PostgreSQL on the same machine,
we stored each relational database on separate SSD drives.
Other than the difference in DRAM space and storage needed for two baselines,
the machines are identical.
The client application running the queries and measuring round-trip time
is connected to the server through a 1GB wired link through
a 10GB back-plane switch, same as both servers.
The client application was implemented using Python 3 for both VDMS and the baselines.
Figure~\ref{fig:systems} shows a logical view of the difference between the
interaction of the client application (retrieves metadata and
images) with VDMS (left), the PostgreSQL baseline (middle), and the MySQL baseline (right).

It is worth noting that the images are stored in a shared repository
(ext4 filesystem on a RAID 6 configuration of 16TB) that both
Apache WebServer and VDMS have direct access.
In the case of videos, only the first frame is used for the image search.
% More information and evaluation on video-specific functionalities can be found
% in the appendix of this work.
In the case of the baselines, metadata is
stored in MySQL and PostgreSQL using an attached SSD disk.
Even if VDMS has native support for Optane Persistent Memory,
we do not use it in this experiment because of fairness of
comparison with respect to MySQL and PostgreSQL, which were not designed for
Persistent Memory type of storage.
The benefits of Persistent Memory for metadata and a full evaluation of
the PMGD subsystem is material for another paper,
and outside the scope of this evaluation.
For this experiment, in the case of VDMS we simply use a similar
attached SSD disk to store metadata.
Even if PMGD, the graph database used by VDMS, is designed for persistent memory,
it can deliver good performance when using SSDs directly, while still
providing ACID-compliant transactions.

For the metadata, we built VDMS, MySQL, and PostgreSQL databases
using the YFCC100M dataset with incremental database sizes.
For simplicity, we named the database based on the approximate number of images
it contains, as follows: 1M, 5M, 10M, 50M, 100M.
The databases have comparable number of elements.
The exact number of images/elements in each database are shown in
Table \ref{table:vdmsnodes} and \ref{table:mysqltables}.
The differences can be attributed to minor failures in data
preparation/loading because of incomplete/inconsistent formatting,
which is common in large datasets \cite{failures}.
In our set up, that difference is very small: less than 0.253\% in terms of
number of elements (images and/or metadata information).

\subsubsection{Data Representation}

@@

\textbf{MySQL/Baseline 1:}
Each MySQL database is created in a similar manner as VDMS
but the data is represented as three tables, following the relational model:
1) \textit{images} table: contains one row per image,
and a column for each property
associated with the images (some of which are listed in Section \ref{dataset});
2) \textit{taglist} table: contains one row per autotag element
(always 1,570 rows);
3) \textit{autotags} table: contains one row per autotag
assigned to an image. Each row contains a foreign key to the
image, a foreign key to the tag, and
the probability assigned to that tag belonging to that image.
Given that there are 8 autotags, on average, per image, the \textit{autotags}
table has around 8 times the number of rows present in the
\textit{metadata} table, as can be seen in Table~\ref{table:mysqltables}.
Using a Python client and simple queries, the \textit{taglist}
table is read from the list of tags with an auto-incremented
\textit{tagid} as a primary key, and the metadata table
is read from the YFCC100M metadata using the identifier as a primary key.
The \textit{autotags} table contains the generated autotags and
probabilities for entries of the \textit{images} table.
To generate the table, we split the \textit{autotags} data for each database
by the image identifier and autotag into new files.
The new files are read into the \textit{autotags} table with the image
identifier and \textit{tagid} as foreign keys.

@@

\textbf{PostgreSQL (Baseline 2):}
Each PostgreSQL database is created in the exact same manner as MySQL
where data is represented as three tables:
\textit{images}, \textit{taglist}, and \textit{taglist}.
PostgreSQL works well out-of-box and optimizations were not necessary,
as in MySQL, to load or execute queries.
When executing queries for larger databases,
it is pertinent to have enough disk space for both
the database and tablespace.
The tablespace is associated with a database and stores the temporary files
created within the database object ~\cite{postgresql}.
When processing the larger databases, additional
storage may be needed for the tablespace or a diskfull error may occur.
To avoid this, we allocated a separate SSD disk specifically
for the tablespace of all databases.

>>>>

We have used the YFCC100M dataset to evaluate different aspects of our system.
We use the images in the dataset and its associated metadata to implement
an image-search engine based on properties associated with those images.
This is a very common use-case we have encountered when building
applications such as smart-retail, sports applications, and video summarization.
For these type of applications, the starting point is usually
a large set of data that must be curated
before proceeding with the data processing (such as neural network training).
In order to evaluate the different aspects of the performance on the
image search, we have built different baselines following the methodology
used in the industry, and following what we have done in the past
in order to build an image search engine, which we describe in Section \ref{images}.
% We have also included performance evaluation on the Video and Feature Vectors
% functionality in the appendix for further reference, together with a
% description of the key aspects of those functionalities.

@@

The Yahoo! Flickr Creative Commons 100m (YFCC100M) dataset is a large
collection of 100 million public Flickr media objects created to provide free,
shareable multimedia data for research.
This dataset contains approximately 99.2 million images and 0.8 million videos
with metadata characterized by 25 fields such as the unique identifier, userid,
date the media was taken/uploaded, location in longitude/latitude coordinates,
device type the media was captured, URL to download the media object,
and the Creative Commons license type information.
The YFCC100M dataset also contains \textit{autotags}
provided as a set of comma-separated concepts such as people, scenery, objects,
and animals from 1,570 trained machine learning classifiers ~\cite{Thomee_2016}.
Together with each \textit{autotags}, there is a probability associated with
each tag to indicate certainty of the classification.
This is, an image can have the \textit{autotags} "people", "person", "party",
"outdoor", and each \textit{autotag} assigned will be accompanied by a
probability of that \textit{autotags} being present in that image/frame.
% We have also used feature vectors generated for every image and first frame
% of every video \cite{features} to implement a similarity search.
Given that there is no standard benchmark oriented towards visual data queries,
we have built a series of queries to filter this dataset that is modeled after
our internal use cases for many of the mentioned applications we have worked
with.

@@

Given that there are no other open-source systems that provide similar
functionality and interfaces as VDMS (i.e., transactionally dealing with images
and metadata behind a single interface), we have implemented two equivalent
visual data management system as baselines,
desing specifically for the image search use case,
and comprised of a combination of widely available, off-the-shelf components.
The first baseline uses MySQL Server 5.7 (for storing metadata),
Apache Web Server 2.4.18 (as interface for image access), and
OpenCV 3.3 (to provide pre-processing operations on images).
We chose Apache Web Server to work as a file server (serving the image files),
because it provides the lowest possible overhead
(when compared to other object store system), at the expenses of providing less
functionality (authentication, data integrity checking, etc).
The other baseline system replaces the MySQL Server 5.7 with a
most advanced open-source relational database, PostgreSQL 9.5 \cite{postgresql}.
We decided to use relational databases in the baseline systems instead of
non-relational databases because of their maturity, and also because it is
the most common tools used for storing and querying metadata
~\cite{Jatana2012, li_2019}:

\begin{itemize}
    \item Relational databases support atomicity, consistency, isolation,
    and durability (ACID) while non-relational may compromise some ACID properties.
    \item The YFCC100m data is clearly structured, and can be modeled with relatively ease with a relational database.
    \item We need to efficiently collate and return metadata records,
    and SQL engines are very mature and optimized for that task.
\end{itemize}

The baseline implementations only partially replicate the functionalities
that VDMS offers when it comes to image and metadata handling, and it was
built for the purpose of an ad-hoc image search implementation.
The baseline implementations are based upon internal tools used for ML-based
pipelines for visual data, which is common practice
in the industry \cite{haystack, tao}.
We have implemented a set of client-side applications that take care
of retrieving the components from the different systems, and applies
pre-processing operations when needed.

For all our experiments, we use two servers with Ubuntu 20.04,
one hosting a VDMS server and another hosting the baseline implementations.
Both servers have a dual-socket Intel\textsuperscript{\textregistered}
Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
each CPU with 28 physical cores with hyper-threading enabled,
for a total of 112 logical cores per server.
The server hosting MySQL and PostgreSQL has 256GB of DDR4 DRAM,
while the server hosting VDMS has 194GB of DDR4 DRAM.
We decided to run the VDMS server in the machine with less DRAM to make
sure MySQL and PostgreSQL had no disadvantage. Previous evaluation
indicated smaller footprint in the case of VDMS when
compared to similar baselines based on MySQL.
To build both MySQL and PostgreSQL on the same machine,
we stored each database on separate SSD drives.
Other than the difference in DRAM space and storage needed for two baselines,
the machines are identical.
The client application running the queries and measuring round-trip time
is connected to the server through a 1GB wired link through
a 10GB back-plane switch, same as both servers.
The client application was implemented using Python 3 for both VDMS and the baselines.
Figure~\ref{fig:systems} shows a logical view of the difference between the
interaction of the client application (retrieves metadata and
images) with VDMS (left), the PostgreSQL baseline (middle), and the MySQL baseline (right).

It is worth noting that the images are stored in a shared repository
(ext4 filesystem on a RAID 6 configuration of 16TB) that both
Apache WebServer and VDMS have direct access.
In the case of videos, only the first frame is used for the image search.
% More information and evaluation on video-specific functionalities can be found
% in the appendix of this work.
In the case of the baselines, metadata is
stored in MySQL and PostgreSQL using an attached SSD disk.
Even if VDMS has native support for Optane Persistent Memory,
we do not use it in this experiment because of fairness of
comparison with respect to MySQL and PostgreSQL, which were not designed for
Persistent Memory type of storage.
The benefits of Persistent Memory for metadata and a full evaluation of
the PMGD subsystem is material for another paper,
and outside the scope of this evaluation.
For this experiment, in the case of VDMS we simply use a similar
attached SSD disk to store metadata.
Even if PMGD, the graph database used by VDMS, is designed for persistent memory,
it can deliver good performance when using SSDs directly.

For the metadata, we built VDMS, MySQL, and PostgreSQL databases using
the YFCC100M dataset with incremental database sizes to study the effects at scale.
For simplicity, we named the database based on the approximate number of images
it contains, as follows: 1M, 5M, 10M, 50M, 100M.
The baseline systems have comparable number of elements.
The exact number of images/elements in each database are shown in
Table \ref{table:vdmsnodes} and \ref{table:mysqltables}.
The differences can be attributed to minor failures in data
preparation/loading because of incomplete/inconsistent formatting,
which is common in large datasets \cite{failures}.
In our set up, that difference is very small: less than 0.253\% in terms of
number of elements (images and/or metadata information).

\subsubsection{Data Representation}
We detail the data representation for each of the systems we implemented:

@@

\textbf{MySQL-Based System:}
Each MySQL database is created in a similar manner as VDMS
but the data is represented as three tables, following the relational model:
1) \textit{images} table: contains one row per image,
and a column for each property
associated with the images (some of which are listed in Section \ref{dataset});
2) \textit{taglist} table: contains one row per autotag element
(always 1,570 rows);
3) \textit{autotags} table: contains one row per autotag
assigned to an image. Each row contains a foreign key to the
image, a foreign key to the tag, and
the probability assigned to that tag belonging to that image.
Given that there are 8 autotags, on average, per image, the \textit{autotags}
table has around 8 times the number of rows present in the
\textit{metadata} table, as can be seen in Table~\ref{table:mysqltables}.
Using a Python client and simple queries, the \textit{taglist}
table is read from the list of tags with an auto-incremented
\textit{tagid} as a primary key, and the metadata table
is read from the YFCC100M metadata using the identifier as a primary key.
The \textit{autotags} table contains the generated autotags and
probabilities for entries of the \textit{images} table.
To generate the table, we split the \textit{autotags} data for each database
by the image identifier and autotag into new files.
The new files are read into the \textit{autotags} table with the image
identifier and \textit{tagid} as foreign keys.

@@

\textbf{PostgreSQL-Based System:}
Each PostgreSQL database is created in the exact same manner as MySQL
where data is represented as three tables:
\textit{images}, \textit{taglist}, and \textit{taglist}.
PostgreSQL works well out-of-box and optimizations were not necessary,
as in MySQL, to load or execute queries.
When executing queries for larger databases,
it is pertinent to have enough disk space for both
the database and tablespace.
The tablespace is associated with a database and stores the temporary files
created within the database object ~\cite{postgresql}.
When processing the larger databases, additional
storage may be needed for the tablespace or a diskfull error may occur.
To avoid this, we allocated a separate SSD disk specifically
for the tablespace of all databases.
